Approximation as Differentiation

### 2025-06-28 main/first-order.tex
\Lift macro.

To interpret the syntactic type constructor $\synVar{L}$, consider the strong monad $(\mathcal{L}, \eta, \mu)$
which acts on a poset $X$ to extend it with a distinguished bottom element $\bot$ and which acts on a monotone
function $f$ between posets to extend it with a mapping from $\bot$ to $\bot$. $\mathcal{L}$ is also a strong
monad on meet-semilattices. On (bounded) join-semilattices, $\mathcal{L}$ induces a costrong comonad
$(\mathcal{L}, \varepsilon, \delta)$, with the bottom element of $X$ needed to implement the counit
$\varepsilon_X: \mathcal{L}(X) \to X$. These two combine to form a strong lifting monad in $\LatGal$ with unit
$\varepsilon_X \dashv \eta_X$ and multiplication $\delta_X \dashv \mu_X$.

The lifting monad $\mathcal{L}$ in $\LatGal$ is preserved into $\Fam(\LatGal)$. The families construction,
regarded as a covariant 2-functor $\Fam(-): \Cat \to \Cat$, sends functors $F: \cat{C} \to \cat{D}$ to
functors $\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that reassign the target category by postcomposing with
$F$, sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and morphisms $(f, \partial f)$ to $(f, F
\comp \partial f)$, where $F \comp \partial f$ denotes the natural transformation where $(F \comp \partial
f)(x) = F(\partial f(x))$. If $F$ happens to be a (strong) monad on $\cat{C}$, then $\Fam(F)$ is a (strong)
monad on $\Fam(\cat{C})$, computed by extending the original monad $F: \cat{C} \to \cat{C}$ componentwise to
families.

>>>>

To interpret the syntactic type constructor $\synVar{L}$, consider the strong monad $(\Lift, \eta, \mu)$ which
acts on a poset $X$ to extend it with a distinguished bottom element $\bot$ and which acts on a monotone
function $f$ between posets to extend it with a mapping from $\bot$ to $\bot$. $\Lift$ is also a strong monad
on meet-semilattices. On (bounded) join-semilattices, $\Lift$ induces a costrong comonad $(\Lift, \varepsilon,
\delta)$, with the bottom element of $X$ needed to implement the counit $\varepsilon_X: \Lift(X) \to X$. These
two combine to form a strong lifting monad in $\LatGal$ with unit $\varepsilon_X \dashv \eta_X$ and
multiplication $\delta_X \dashv \mu_X$.

The lifting monad $\Lift$ in $\LatGal$ is preserved into $\Fam(\LatGal)$. The families construction, regarded
as a covariant 2-functor $\Fam(-): \Cat \to \Cat$, sends functors $F: \cat{C} \to \cat{D}$ to functors
$\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that reassign the target category by postcomposing with $F$,
sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and morphisms $(f, \partial f)$ to $(f, F \comp
\partial f)$, where $F \comp \partial f$ denotes the natural transformation where $(F \comp \partial f)(x) =
F(\partial f(x))$. If $F$ happens to be a (strong) monad on $\cat{C}$, then $\Fam(F)$ is a (strong) monad on
$\Fam(\cat{C})$, computed by extending the original monad $F: \cat{C} \to \cat{C}$ componentwise to families.
